Carbonate chemistry at Ny-Ålesund

### 2021-11-18 awipev-CO2.Rmd
rewrote ggreg and revised code accordingly

#################### Regression function
# function regression plot with equation (MA or OLS) in title
ggreg <- function (xdata, ydata, type = "OLS", plot = TRUE, title = "title", subtitle = "subtitle", xlab = "x", ylab = "y", proplot = TRUE) 
  { # x and y are the names of the variables
  fit_data <- tibble(xdata, ydata)
  colnames(fit_data) = c("xdata", "ydata")
  fit <- print(lmodel2::lmodel2(ydata ~ xdata))
if (type == "OLS") ty = 1 #one selects OLS only
if (type == "MA") ty = 2 #one selects MA only
#reg <- fit$regression.results[ty,] #one selects OLS only
intercept <- fit$regression.results$Intercept[ty]
slope <- fit$regression.results$Slope[ty]
predicted <- fit$regression.results$Intercept[1] + xdata * fit$regression.results$Slope[1]
error <- xdata - predicted
if (plot == TRUE) {
  plt <- ggplot(data = fit_data, aes(x = xdata, y = ydata)) + 
  geom_point(na.rm = TRUE, size = 2, colour = "blue") +
  geom_abline(aes(intercept = intercept, slope = slope),
              colour = "blue")  + 
  labs(title = paste(title, 
                     "\ntype: ", type,
                     "; R2 = ", round(fit$rsquare, digits = 2),
                     "; Intercept =", signif(intercept, 3),
                     "; Slope =", signif(slope, 3),
                     "\nP =", signif(fit$P.param, 3),
                    "rmse =", signif(sqrt(mean(error^2, na.rm=TRUE)), 3)),
       subtitle = subtitle,
       x = xlab,
       y = ylab) +
    theme_bw() +
    theme(plot.title = element_text(size = 7))
    if (proplot == TRUE) { # property-property plot
      plt <- plt +
        geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
        theme(aspect.ratio = 1)
      }
  return(plt)
  } #else {
  #return(fit)
#}
}

@@

d_hour <- d_hour %>%
  dplyr::filter(datetime <= as.POSIXct("2021-01-10")) %>% #cutoff date
  dplyr::select(-pco2_inst) %>% # this is the serial number of pCO2 sensor, not needed
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
# read a file provided by Philipp on 2021-04-09 which fills most gaps of pressure data
# depth_complete replaces all other depth and pressure variables
svl_remos_depth <- tibble(
  readRDS(file = paste0(
    path,
    "fb_data/svl_remos_depth.rds"
  )))
d_hour <- dplyr::left_join(d_hour, svl_remos_depth, by = 'datetime') %>% 
  dplyr::select(-adcp_fill, -ctd_fill) %>% 
  dplyr::rename(depth = depth_complete)
# check_plot <- d_hour %>% 
#   ggplot() + geom_point(aes(x=datetime, y=depth, col='blue', alpha = 0.7)) +
#   geom_point(aes(x=datetime, y=depth, col='red', alpha = 0.3))

@@

@@

```{r measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
title <- "pCO2 normalized at in situ temperature"
p <- ggreg(
  xdata = d_all$pco2_fb_insi_temp,
  ydata = d_all$pco2_calc_fb_insi_temp,
  type = "MA",
  plot = TRUE,
  title = title,
  subtitle = "",
  xlab = "Measured pCO2",
  ylab = "Calculated pCO2",
  proplot = TRUE
)
print(p)
ggsave(filename = "figures/pco2_calc_meas.png", plot = p)
```

The relationship between the measured (Contros) and calculated pCO2 (blue line) is relatively poor. It is far from the 1:1 relationship (black dashed line). The slope is `r signif(fit$regression.results[2,3], 3)`) and its 95% confidence interval does include 1; r2 is `r round(fit$rsquare, digits=2)`.

**Figure legend**: pCO2 calculated from AT and CT as a function of pCO2 measured using a Contros sensor. All data are normalized at in situ temperature. The black dotted line is the 1:1 line while the blue solid line shows the line obtained by a major axis regression (Legendre, 2018- Pierre Legendre (2018). lmodel2: Model II Regression. R package version 1.7-3. https://CRAN.R-project.org/package=lmodel2
<!-- ======= -->
<!--   theme(aspect.ratio=1,plot.title = element_text(size=7)) +  -->
<!--   coord_fixed(ratio = 1 ,xlim=c(200,460) , ylim=c(200,460))+ -->
<!--   labs(x="Contros pCO2", y="Calculated pCO2") + theme_bw()   -->
<!-- ``` -->

The relationship between the Contros and calculated pCO2 is not good at all in the sense that it is far from the 1:1 line, despite a good r2 (`r round(fit$rsquare, digits=2)`).
<!-- >>>>>>> cb6a0f5891d0a4e43735bed5507c62760ae133cb -->

```{r plot measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
# print(p)
# 
```

@@

            ylab = "pH @ 4°C calculated (pCO2+AT) with Lueker") +

@@

@@

            ylab = "pH @ 4°C calculated (pCO2+AT) with Sulpis")  +

@@

            ylab = "pH @ 4°C calculated (AT+CT) with Lueker") +

@@

            ylab = "pH @ 4°C calculated (AT+CT) with Sulpis")  +

>>>>

#################### Regression function
# function regression plot with equation (MA or OLS) in title
ggreg <- function (xdata, ydata, type = "OLS", plot = TRUE, title = "title", subtitle = "subtitle", xlab = "x", ylab = "y", proplot = TRUE, ...) 
  { # x and y are the names of the variables
  fit_data <- tibble(xdata, ydata)
  colnames(fit_data) = c("xdata", "ydata")
  fit <- print(lmodel2::lmodel2(ydata ~ xdata))
if (type == "OLS") ty = 1 #one selects OLS only
if (type == "MA") ty = 2 #one selects MA only
#reg <- fit$regression.results[ty,] #one selects OLS only
intercept <- fit$regression.results$Intercept[ty]
slope <- fit$regression.results$Slope[ty]
predicted <- fit$regression.results$Intercept[1] + xdata * fit$regression.results$Slope[1]
error <- xdata - predicted
if (plot == TRUE) {
  plt <- ggplot(data = fit_data, aes(x = xdata, y = ydata)) + 
  geom_point(na.rm = TRUE, size = 2, colour = "blue") +
  geom_abline(aes(intercept = intercept, slope = slope),
              colour = "blue")  + 
  labs(title = paste(title, 
                     "\ntype: ", type,
                     "; R2 = ", round(fit$rsquare, digits = 2),
                     "; Intercept =", signif(intercept, 3),
                     "; Slope =", signif(slope, 3),
                     "\nP =", signif(fit$P.param, 3),
                    "rmse =", signif(sqrt(mean(error^2, na.rm=TRUE)), 3)),
       subtitle = subtitle,
       x = xlab,
       y = ylab) +
#    lims(x = c(min(fit_data$xdata), max(fit_data$xdata)), y = c(min(fit_data$ydata), max(fit_data$ydata))) +
    theme_bw() +
    theme(plot.title = element_text(size = 7))
    if (proplot == TRUE) { # property-property plot
      plt <- plt +
        geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
        theme(aspect.ratio = 1)
    }
  return(list(plot = plt, fit = fit))
  } else {
  return(fit)
    }
}

@@

d_hour <- d_hour %>%
  dplyr::filter(datetime <= as.POSIXct("2021-01-10")) %>% #cutoff date
  dplyr::select(-pco2_inst) %>% # this is the serial number of pCO2 sensor, not needed
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  # data post 2020-03-01 eliminated because no maintenance and lots of algae
  dplyr::mutate(pco2_fb = ifelse(datetime >= as.POSIXct("2020-03-01 00:00"), NA, pco2_fb)) %>%
  arrange(datetime)
# read a file provided by Philipp on 2021-04-09 which fills most gaps of pressure data
# depth_complete replaces all other depth and pressure variables
svl_remos_depth <- tibble(
  readRDS(file = paste0(
    path,
    "fb_data/svl_remos_depth.rds"
  )))
d_hour <- dplyr::left_join(d_hour, svl_remos_depth, by = 'datetime') %>% 
  dplyr::select(-adcp_fill, -ctd_fill) %>% 
  dplyr::rename(depth = depth_complete)
# check_plot <- d_hour %>% 
#   ggplot() + geom_point(aes(x=datetime, y=depth, col='blue', alpha = 0.7)) +
#   geom_point(aes(x=datetime, y=depth, col='red', alpha = 0.3))

@@

**Note** that the Contros pCO2 values after 2020-03-01 were not used because the Covid-19 pandemic prevented maintenance and the setup of a freshly calibrated sensor. As the result, algae becalme inceasingly abundant, pulling pCO2 down and further away from values calculated from CT and AT.

@@

```{r measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
title <- "pCO2 normalized at in situ temperature"
reg <- ggreg(
  xdata = d_all$pco2_fb_insi_temp,
  ydata = d_all$pco2_calc_fb_insi_temp,
  type = "MA",
  plot = TRUE,
  title = title,
  subtitle = "",
  xlab = "Measured pCO2",
  ylab = "Calculated pCO2",
  proplot = TRUE
)
p <- reg$plot
ggsave(filename = "figures/pco2_calc_meas.png", plot = p)
```

The relationship between the measured (Contros) and calculated pCO2 (blue line) is relatively poor. It is far from the 1:1 relationship (black dashed line). The slope is `r signif(reg$fit$regression.results[2,3], 3)`) and its 95% confidence interval does include 1; r2 is `r round(reg$fit$rsquare, digits=2)`.

**Figure legend**: pCO2 calculated from AT and CT as a function of pCO2 measured using a Contros sensor. All data are normalized at in situ temperature. The black dotted line is the 1:1 line while the blue solid line shows the line obtained by a major axis regression (Legendre, 2018- Pierre Legendre (2018). lmodel2: Model II Regression. R package version 1.7-3. https://CRAN.R-project.org/package=lmodel2

```{r plot measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
print(p$plot)
# 
```

@@

            ylab = "pH @ 4°C calculated (pCO2+AT) with Lueker", 
            proplot = TRUE)$plot +

@@

        axis.text = element_text(size = 10))

@@

            ylab = "pH @ 4°C calculated (pCO2+AT) with Sulpis",
            proplot = TRUE)$plot +

@@

            ylab = "pH @ 4°C calculated (AT+CT) with Lueker",
            proplot = TRUE)$plot +

@@

            ylab = "pH @ 4°C calculated (AT+CT) with Sulpis",
            proplot = TRUE)$plot  +
